### 2019-03-29 5 - FSCD 2019 - Submitted/Reviews/rebuttal v1.txt
reviews, rebuttal, reorganise

Review 1:

(A) Why isn't there any **leap** edge?

The purpose of a leap edge is to explicitly record the dependency between quantifiers: when the eigenvariable (x) of a universal quantifier is used in the witnessing term (t{x}) to an existential quantifier. This corresponds to our "dependency" relation (Definition 10). The relevant part can be paraphrased: if (x) is the eigenvariable for a quantification "forall x.Y", and an explicit substitution [t{x}/y] ranges over a formula X, then there is a dependence (a "leap" or "jump") between Y and X.

We do not need leap edges for coalescence; we believe that likewise, contractibility for MLL1 unification nets would not need them.

(B) Is there a mismatch between slicing and unification?

Yes. For geometric correctness, it seems essential to reconstruct the dependencies (or "leaps", or "jumps") globally. In MLL1, where there are no slices, a single global unifier (for all links) can be constructed, from which the dependency is then derived.

With slices, the situation is much more involved. Dependencies (leaps) must be made explicit between existential quantifiers and products (see "Expansion", pages 8--9) as well as universal quantifiers. These interact in highly intricate ways, which the unifying coalescence algorithm resolves incrementally and locally. To do so globally, as a geometric criterion would, is a huge combinatorial challenge, beyond the scope of this paper.

==============================

Review 2:

(C) Is the equivalence generated by unification a form of observational equivalence?

We believe, yes, but none of us is an expert.

(D) Both proofs and witness nets are not considered up to alpha-equivalence.

In fact, we may alpha-convert at the level of proofs and proof nets; but for the sanity of the translations both ways, not at the level of individual formulas in a sequent proof.

Normally, indeed, every universal instantiation must have a unique eigenvariable. But for our proof nets, we only need (and have) this property for each slice individually.

==============================

Reviews 2 and 3:

(E) Can this be extended to MALL1?

The situation is mixed. There are two ways forward that are likely to succeed:
 - combine witness nets and MALL slice nets [Hughes & Van
   Glabbeek 2005], using a slicing criterion;
 - combine unification nets and MALL conflict nets [Hughes &
   Heijltjes 2016], using a coalescence criterion.
For both, there are still significant combinatorial challenges (e.g. because coalescence for MALL is not a straightforward extension of that for ALL). Other combinations are unlikely to work, or the groundwork is still incomplete: slice nets do not support coalescence (this looks fundamentally problematic), and conflict nets do not yet have a slicing criterion (work in progress).