### 2019-04-01 5 - FSCD 2019 - Submitted/Reviews/rebuttal v2.txt
rebuttal v2

Review 1 is overall very positive but makes the vote for acceptance
dependent on two technical questions that we address in (A) and (B)
below.

Review 2 and 3 both ask about MALL1, which we discuss in (C) below.

Further minor questions by Review 2 and 3 are addressed in (D), (E),
and (F).

=====================================
Review 1:

(A) Why isn't there any **leap** edge in the slicing criterion?

There are! However, they are not explicit. They occur in the form of a
"dependency" relation, as part of the geometric correctness
criterion. (This is standard for graphical first-order proofs; see
e.g. Miller's Expansion Tree Proofs.)

The purpose of a leap edge in other proof net like presentations is to
explicitly record the dependency between quantifiers: when the
eigenvariable (x) of a universal quantifier is used in the witnessing
term (t{x}) to an existential quantifier.

We capture this in our "dependency" relation (Definition 10). The
relevant part can be paraphrased: if (x) is the eigenvariable for a
quantification "forall x.Y", and an explicit substitution [t{x}/y]
ranges over a formula X, then there is a dependence (i.e. a "leap", or
a "jump") between Y and X.

Just like the "leaps" in Hughes's MLL1 nets, our dependency is not
explicit, but derived from the data structure (from the occurrences of
eigenvariables in the explicit substitutions). But note that in MLL1,
the derived leaps are explicitly part of the "switching graphs" of the
correctness criterion, so that they can interact with tensors to
create cycles. In ALL1 there are no tensors to interact with leaps, so
we can have a completely separate “dependency acyclicity” condition
(our "dependency-correct" condition).

In contrast with geometric correctness, we do not need leap edges for
coalescence; we believe that likewise, contractibility for MLL1
unification nets would not need them.

(B) Is there a mismatch between slicing and unification?

Yes. For geometric correctness, it seems essential to reconstruct
dependencies globally. In MLL1, where there are no slices, a single
global unifier (for all links) can be constructed, from which the
dependency is then derived (as "leap" edges).

With slices, the situation is much more involved. Dependencies must be
made explicit between existential quantifiers and products (see
"Expansion", pages 8--9) as well as existential and universal
quantifiers. These interact in highly intricate ways - which the
unifying coalescence algorithm resolves incrementally and locally. To
do so globally, as would be necessary for a geometric criterion, is a
huge combinatorial challenge, beyond the scope of this paper.

=====================================
Review 2 and 3:

(C) Can this be extended to MALL1?

The situation is mixed. There are two ways forward that could succeed:

 - combine witness nets and MALL slice nets [Hughes & Van
   Glabbeek 2005], using a slicing criterion;

- combine unification nets and MALL conflict nets [Hughes &
   Heijltjes 2016], using a coalescence criterion.

For both, there are still significant combinatorial challenges
(e.g. because coalescence for MALL is not a straightforward extension
of that for ALL). Other combinations are unlikely to work, or the
groundwork is still incomplete: MALL slice nets do not support
coalescence (this looks fundamentally problematic), and MALL conflict
nets do not yet have a slicing criterion (work in progress).

=====================================
Review 2:

(D) Is the equivalence generated by unification a form of observational equivalence?

We believe, yes, but none of us is an expert.

(E) Both proofs and witness nets are not considered up to alpha-equivalence.

In fact, we may alpha-convert at the level of proofs and proof nets; but for the sanity of the translations between them, not at the level of individual formulas in a sequent proof.

Normally in proof nets, indeed, every universal instantiation should have a unique eigenvariable. But here we only need this property for each slice individually.

=====================================
Review 3:

(F) Unification seems to break the variable naming convention.

Indeed - the standard unification algorithm would. But we can stipulate explicitly that an "initial witness map" (the MGU for a link) uses fresh variables in its range, distinct from any existentially quantified ones (which occur in its domain); then all is well.